\textbf{Title}

### 2015-08-07 paper/draft_3.tex
updated plots

\section{Abstract}
\section{Introduction}
In recent years, various shrinkage and thresholding procedures have become ever more popular in many applications where sparsity is desired. Here ``sparsity'' is loosely defined in that it also includes the case where many terms are near-zero, as opposed to exactly zero. The most prominent usage of such procedures appear in multiresolution denoising analyses such as wavelets, where sparsity in a transformed domain often corresponds to smoothness in the data space [references]. Other examples include estimation of covariance and precision matrices (though thresholding is used more often here since zero terms are actually desired) [references], accounting for different measurement precisions in FDR control for multiple testing [references], or more generally optimizing penalized likelihood problems such as LASSO [references]. In this paper we will look at a flexible and adaptive shrinkage method named ASH, and illustrate its value in two key areas of univariate wavelet denoising.\bigskip\\
Before delving deeper into the two denoising tasks, we first present a brief overview of wavelet estimators. These estimators have become extremely popular in nonparametric regression problems following the seminal paper by Donoho \& Johnstone (1994). Wavelet methods are locally adaptive in the sense that they achieve the optimal minimax rate over a wide variety of functions, yet are also computationally faster than many other adaptive methods such as variable-bandwidth kernel methods [antoniadis 2001]. While classical thresholding estimators have been shown to possess optimal asymptotic properties, an extensive simulation study by Antoniadis et al. (2001) has demonstrated that various Bayesian shrinkage methods are usually more data adaptive than thresholding methods in finite samples, outperforming the latter in terms of mean squared error (MSE). However, many Bayesian methods require specific distributional assumptions, and are not easily extensible to more complex problems, two of which we subsequently describe. On the other hand, our method simplies the task by making approximations, while still possessing good finite sample performance and computational efficiency. These approximations also allows us to use ASH as the common shrinkage procedure in both the problems.\bigskip\\
The first problem relates to mean and variance estimation with Gaussian noise. Although the framework for homoskedastic Gaussian errors has been thoroughly developed in the wavelet literature [references], existing methods dealing with heteroskedastic errors have been far and few between. Fan \& Yao (1998) estimated the variance by smoothing the squared residuals using local polynomial smoothing, while Brown \& Levine (2007) employed difference-based kernel estimators. Making use of the local adaptivity of wavelet methods, Cai \& Wang (2008) improved upon previous variance estimation methods using a wavelet thresholding approach on first order differences. Here we present an approach that estimates both the mean and variance accurately by incorporating ASH into a novel wavelet denoising framework. [mention the lack of available software?]\bigskip\\
The second task is to denoise a Poisson distributed signal. This often occurs in the experimental sciences such as gamma-ray burst signals in astronomy [references], and more recently in genetics where high throughput sequencing data is of interest. This problem is interesting in that there are many different ways to perform signal recovery. Variance stabilizing techniques together with normal approximations have been proposed by Donoho (1993) and Fryzlewicz \& Nason (2001), by exploiting the mean-variance relationship of a Poisson distribution. Kolaczyk (1997, 1999a) derived thresholds achieving optimal asymptotic properties in the context of wavelet transformations, similar to the thresholds in the i.i.d. Gaussian case.  However, variance stabilizing methods are computationally inefficient due to the presence of external cycle-spinning, and threshold-based methods may not result in satisfactory finite sample performance. Furthermore, both these methods are quite sensitive to the choice of the primary resolution level as simulations have shown. Multiscale analysis using recursive dyadic partitions within a Bayesian framework was later developed by Kolaczyk (1999b) to make use of a particular form of likelihood factorization, but a relatively inflexible conjugate prior was chosen. We will improve upon the prior (and likelihood) specification by using ASH as the main shrinkage procedure in a Bayesian framework similar to that of Kolaczyk (1999b). \bigskip\\
Both the aforementioned problems are harder to deal with than the classical problem of i.i.d. Gaussian errors, and usually requires one to extend the ideas from the latter task. Before proceeding to describe the two problems in detail, we will first describe briefly the portion of ASH that performs shrinkage. \bigskip\\
As a generic shrinkage method, ASH takes as input a vector of estimates $\hat{\Gb}_i,i=1,...,n$ and their standard errors $\hat{\s}_i\equiv se(\hat{\Gb}_i),i=1,...n$, and outputs shrunken estimates of $\bm{\Gb}$. Specifically, the likelihoods for the true parameters $\Gb_i,i=1,...n$ are assumed to be iid $N(\Gb_i;\hat{\Gb}_i,\hat{\s}_i^2)$. ASH then assumes exchangeability of the $\Gb_i$'s and sets the following ``shrinkage'' prior on $\Gb_i$ for each $i$:
\Gb_i|\bm{\pi}=\sum_{k=1}^m \pi_k N(\Gb_i;0,s_k^2)
Note that $\bm{\pi}$ and $\bm{s}^2$ are shared across all the $\Gb_i$'s, allowing one to borrow information across all observations. The full likelihood for $\bm{\pi}$ given $\bm{\hat{\Gb}}$ and $\bm{\hat{\s}}$ is then a product of likelihoods for $\bm{\pi}$ given each $\hat{\Gb}_i$ and $\hat{\s}_i$. Since the likelihoods are Gaussian and the priors are mixtures of Gaussians, the posterior distribution of $\Gb_i$ given $\hat{\Gb}_i$ and $\bm{\hat{\pi}}$ is analytically tractable, where the mixing proportions $\bm{\hat{\pi}}$ are maximum likelihood estimates. The posterior mean of $\Gb_i$ is taken to be the point estimate of $\Gb_i$, resulting in a shrinkage procedure.\bigskip\\
The flexibility and computational efficiency of ASH is immediately obvious, which motivates us to extend this method to the more complex wavelet settings described here. As an Empirical Bayes procedure, ASH borrows information across all available data and is extremely data adaptive. At the same time, the key advantage of ASH over other Empirical Bayes procedures such as EbayesThresh [Johnstone \& Silverman (2005)] and the commonly used ``spike-and-slab'' prior (is there a name for this?) variable variances are allowed as input, a feature which latter methods lack. This is also the primary reason we can use ASH as the only shrinkage procedure in both the problems described above. Furthermore, Gaussian mixture priors can effectively approximate any unimodal distribution, allowing for a more flexible prior than many other procedures. In tackling these two specific problems, we can speed up the entire signal denoising problem via likelihood approximations (discussed in the next section), allowing our method to be computationally efficient without compromising accuracy. Typical wavelet methods require the specification of a primary resolution level as a ``tuning'' parameter, which could substantially influence the accuracy of the method. However, applying ASH to every resolution in the wavelet transformationwe spares us the need for such a parameter. For an in-depth understanding of the original motivations and applications of ASH, see Stephens (??). With the main shrinkage method accounted for, we will proceed to explore the two aforementioned problems in detail. Note that one only needs to supply a vector of estimates $\hat{\Gb}$ and their standard errors se$(\hat{\Gb})$ to ASH to obtain posterior estimates; hence, we will focus on obtaining these estimates and standard errors when describing our methods.
\section{Method}
\subsection{Gaussian denoising with heteroskedastic errors}
Consider the nonparametric regression model with i.i.d. Gaussian errors:
\begin{eqnarray}\label{eq:1d gaussian model}
Y_i=\mu_i+\Ge_i
for $i=1,...,n$, where $\bm{Y}$ is the vector of observations, $\bm{\mu}$ is the mean curve, and $\Ge_i$ are independent $N(0,\s^2)$ noise. Assume that $n=2^J$ for some integer $J$, as is standard in the wavelet literature. To motivate our method, we first describe wavelet shrinkage from a Bayesian perspective.

The wavelet coefficients $\bm{d}$ are given by
\bm{d}=W\bm{Y}
\end{eqnarray
where $W$ denotes the orthogonal $n\times n$ matrix corresponding to the orthonormal wavelet basis chosen.
Since $\mathbb{V}(Y)=\s^2I$,
\begin{eqnarray}\label{eq:waveletcoef}
\bm{d}\sim N_n(\bm{\Ga},\s^2I)
where $\bm{\Ga}=W\mathbb{E}(\bm{Y})$. This implies that the likelihood for $\bm{\Ga}$ factorizes into a product of likelihoods for $\Ga_i$, $i=1,...,n$. As such, a natural and computationally convenient approach is to set independent priors on $\Ga_i$ as a form of ``shrinkage''. Here we apply the priors from ASH to each resolution level separately, resulting in
%Although there are many choices for the prior, we use the Gaussian mixture prior from ASH in our method, given by
\begin{eqnarray}\label{eq:ashprior}
\Ga_{jk}=\sum_{l=1}^m \pi_l^{(j)} N(0,(s_l^{(j)})^2)
with $\sum_l \pi_l^{(j)}=1$, for $j=1,...,J$. $\pi_i^{(j)}$ and $(s_i^{(j)})^2$ are hyperparameters that are shared between coefficients in the same resolution level, and $m$ is the number of mixture components. In ASH, $m$ is usually chosen so that it is large enough to reasonably approximate any unimodal distribution, while maintaining an acceptable computational speed. Note that, for notational convenience, we have switched to a double index following standard wavelet convention for indices: here $j=1,...,J$ is the resolution level, and $k=0,...,2^j-1$ is the location within each resolution level $j$. By applying the inverse wavelet transform on the posterior mean $\hat{\bm{\Ga}}$ that ASH produces, one can obtain the posterior mean of $\bm{\mu}$, which serves as an estimate that minimizes the MSE. One could also construct credible bands using the posterior variances.\bigskip\\
While many existing methods do a good job of signal denoising in the presence of i.i.d. (Gaussian) errors, heteroskedastic (but still independent) errors present a different challenge. One key obstacle when dealing with heteroskedastic errors is that the likelihood for $\bm{\Ga}$ does not necessarily factorize. Given the true variances and hence the likelihood, one could compute the posterior distribution using some specified prior. However, this would be computationally cumbersome, and may not be feasible when extended to multiple signals or higher dimensions. A suitable prior might also be difficult to find. As such, one key aspect of our approach is to treat the wavelet coefficients as if they were independent, so that the true likelihood is approximated by a composite likelihood (see Silverman (1999) for an example where this is done).\bigskip\\
To describe our approach in more detail, first assume that the true variance function is known and given by $\mathbb{V}(\Ge_i)=\s_i$, and that the true mean function is given by $\bm{\mu}=(\mu_1,...,\mu_t)$, where $T=2^J$ for some positive integer $J$. Hence
\bm{Y}\sim N_n(\bm{\mu},\Sigma)
where $\Sigma=\textrm{diag}(\s_1,...,\s_T)$ is a diagonal matrix. For the rest of this subsection, we will use the non-decimated wavelet transform (NDWT) instead of the standard DWT to achieve translation invariance. This in turn reduces the presence of artifacts, which often occur near discontinuities in the underlying signal (see eg. Coifman \& Donoho (1995)). Now let $W$ be the matrix associated with the NDWT for a given wavelet basis, so that
\bm{d}\sim N_n(\bm{\Ga},\tilde{\Sigma})
where $\bm{d}$ is the vector of detail coefficients, $\bm{\Ga}=W\bm{\mu}$, and $\tilde{\Sigma}=W\Sigma W^T$. By treating the likelihood for $\bm{\Ga}$ as if it were independent, we can write the likelihood as follows (using the double index mentioned above):
\begin{eqnarray}\label{eq:likelihood}
L(\bm{\Ga}|\bm{d})=\prod_{j=0}^J\prod_{k=0}^{T-1}P(d_{jk}|\Ga_{jk})
where $P(d_{jk}|\Ga_{jk})=\phi(d_{jk};\Ga_{jk},\tilde{\Sigma}_{(jk,jk)})$. Note that there are $n$ coefficients at each resolution level instead of $2^j$ for resolution $j$ due to the NDWT. Here $\phi$ denotes the Gaussian density function. Since $\Sigma$ is diagonal, it is easy to see that $\tilde{\Sigma}_{(jk,jk)}=\sum_{i=1}^T \Sigma_{ii}W_{jk,i}^2$. In our method, we use ASH to assign independent priors to $\Ga_{jk}$ as with \eqref{eq:ashprior}. %ASH also allows us to estimate the $\pi_i$'s via a maximum likelihood approach and fix the $\s_i$'s by using a grid of values. This particular choice of prior specification is extremely flexible, being able to approximate any symmetrical zero-centered distribution, as well as allowing for a relatively easy estimation of the hyperparameters through empirical Bayes. See ASH for more details regarding this choice of prior (?)
While ASH produces shrunken estimates of $\Ga_{jk}$ for $j,k>0$, we estimate $\Ga_{00}$ using the corresponding scaling coefficient, which seems intuitively appealing. Finally, we can obtain an estimate of $\mu$ by using the average basis inverse, which is essentially an average of the inverse wavelet transforms for every shift of the data (see Coifman \& Donoho (1995)) since we are using the NDWT. Although the method described here uses a matrix formulation for easier conceptual understanding, the actual NDWT and inverse transform are done through Mallat's pyramid algorithm, which takes only $n\log(n)$ time.\bigskip\\
Although we have assumed that the true variances were known for mean estimation, the problem of variance estimation itself is a non-trivial one. Here we make the reasonable assumption that the variance function is also spatially structured, in addition to the mean function. The approach we took is similar in spirit to that of Cai \& Wang (2008), since we make use of wavelet decomposition as well. However, while they use first order differences, we look at the squared residuals. As such, we need a reasonably accurate estimator of the mean function to form sensible residuals. For clarity of presentation, we will first assume that the true mean function is known. Define
\begin{eqnarray}\label{eq:varobs1}
Z_i^2=(Y_i-\mu_i)^2
to be the ``observations'' for the unknown variance function. Note that $\mathbb{E}(Z_i^2)=\s_i^2$, so that $Z_i^2$ is unbiased for the true variance at point $i$. At this point we can treat this as yet another mean estimation problem, with the added complication that the variance of $Z_i^2$ has a $\chi^2$ distribution squared. To simplify the problem, we approximate this likelihood by a Gaussian likelihood. Of course, better likelihoods and priors could be used that reflect the skewness in the distributions of the ``observations'', but we found that a normal approximation works reasonably well (especially for smoother variance functions), and has the key advantage of being easy and fast to implement. The variance estimation process is then very similar to one for mean estimation as described above, with a few extra details. The full procedure is outlined in Appendix \ref{app:var estimation}.\bigskip\\
Now that we can now estimate $\bm{\mu}$ given $\bm{\s}^2$ and vice versa, a natural procedure for jointly estimating $\bm{\mu}$ and $\bm{\s}^2$ is an iterative one. That is, we estimate $\bm{\mu}$ given some initial estimate of $\bm{\s}^2$, then estimate $\bm{\s}^2$ given the previous estimate of $\bm{\mu}$, and iterate the process. In our simulations, we have found that two cycles are usually sufficient to produce relatively accurate estimates. To start off the algorithmm, an initial estimate would be the square of first order differences between adjacent points, as discussed in previous variance estimation papers (eg Cai \& Wang (2008)). This estimator has the property that it is approximately unbiased for the variance at the two corresponding points, provided that the mean and the variance function is smooth enough. To be more specific, the initial variance estimates are defined by

@@

Next we turn to another popular nonparametric regression problem and demonstrate how we could make use of ASH. Specifically, we also assume an underlying signal $\Gl_i$, $i=1,...,n$ with $n=2^J$ for some positive integer $J$. Further assume that each data point $Y_i$, $i=1,...,n$ is realized from a Poisson distribution with mean $\Gl_i$ ie. $Y_i=\textrm{Pois}(\Gl_i)$. Our goal is to recover the true intensity $\bm{\Gl}$ as accurately as possible, given $Y_k$. To make our approach easier to understand and relate it to wavelet-based methods, we first summarize the data in a recursive manner (see Kolaczyk (1999) and Nowak (1998):

@@

for resolution $j=0,...,J-1$ and location $k=0,...,2^j-1$, so that we are looking at the sum of adjacent pairs of ``points'' as we move to coarser resolutions, where a point is already itself a sum of the data points. Another way to define the $Y$'s is

@@

Whereas Kolaczyk (1999) sets a conjugate shrinkage prior (a mixture of a point mass at 0.5 and a uniform distribution, similar to the ``spike and slab'' prior for the Gaussian case) for the parameters $\frac{\Gl_{j+1,2k}}{\Gl_{jk}}$ and uses a Binomial likelihood, we felt that the two-component mixture prior is too inflexible. As such, we work with the parameters $\Ga_{sl}$ as defined in \eqref{eq:poisson wc} and approximate the logit-Binomial (is there such a thing?) likelihoods by Gaussian likelihoods. Furthermore, these parameters can be more easily connected to the true wavelet coefficients since they consist of (log) differences rather than ratios.\bigskip\\
To describe the approximation process in more detail, we drop the subscripts for notational convenience. One direct approximation of the likelihood for $\Ga$ is the Laplace approximation, which results in the Gaussian likelihood given by $\phi(\Ga;\hat{\Ga}_{MLE},\mathbb{V}(\hat{\Ga}_{MLE}))$. Here $\phi$ is the normal density function, and $\hat{\Ga}_{MLE}$ is the maximum likelihood estimator. We would then estimate $\mathbb{V}(\hat{\Ga}_{MLE})$ by the square of the asymptotic standard deviation for $\hat{\Ga}_{MLE}$. Specifically,
% for some appropriate estimator $\hat{\Ga}$. This choice of parameters for the Gaussian approximation is also appropriate in the sense that it minimizes the Kullback-Leibler divergence between a Gaussian distribution and the likelihood for $\Ga$ (also known as moment matching).\bigskip\\
%We next consider appropriate estimators for the parameters $\Ga$ and $\sqrt{\mathbb{V}(\hat{\Ga})}$. The obvious choice would be the MLE of $\Ga$ and its asymptotic standard error respectively, given by
\hat{\Ga}_{MLE}=\log(S/F)&\ \ \ S=0,...,N\\
se(\hat{\Ga}_{MLE})=\sqrt{SF/N}&\ \ \ S=0,...,N

>>>>

\section{Abstract}
\section{Introduction}
In recent years, various shrinkage and thresholding procedures have become ever more popular in many applications where sparsity is desired. Here ``sparsity'' is loosely defined in that it also includes the case where many terms are near-zero, as opposed to exactly zero. The most prominent usage of such procedures appear in multiresolution denoising analyses such as wavelets, where sparsity in a transformed domain often corresponds to smoothness in the data space [references]. Other examples include estimation of covariance and precision matrices (though thresholding is used more often here since zero terms are actually desired) [references], accounting for different measurement precisions in FDR control for multiple testing [references], or more generally optimizing penalized likelihood problems such as LASSO [references]. In this paper we will look at a flexible and adaptive shrinkage method named ASH, and illustrate its value in two key areas of univariate wavelet denoising.

Before delving deeper into the two denoising tasks, we first present a brief overview of wavelet estimators. These estimators have become extremely popular in nonparametric regression problems following the seminal paper by Donoho \& Johnstone (1994). Wavelet methods are locally adaptive in that they achieve the optimal minimax rate over a wide variety of functions, yet are also computationally faster than many other adaptive methods such as variable-bandwidth kernel methods [antoniadis 2001]. While classical thresholding estimators have been shown to possess optimal asymptotic properties, an extensive simulation study by Antoniadis et al. (2001) has demonstrated that various Bayesian shrinkage methods are usually more adaptive than thresholding methods in finite samples, outperforming the latter in terms of mean squared error (MSE). However, many Bayesian methods require specific distributional assumptions, and are not easily extensible to more complex problems, two of which we subsequently describe. On the other hand, our method simplies the task by making approximations, while still possessing good finite sample performance and computational efficiency. These approximations also allow us to use ASH as the common shrinkage procedure in both the problems.

The first problem relates to mean and variance estimation with Gaussian noise. Although the framework for homoskedastic Gaussian errors has been thoroughly developed in the wavelet literature [references], existing methods dealing with heteroskedastic errors have been far and few between. Fan \& Yao (1998) estimated the variance by smoothing the squared residuals using local polynomial smoothing, while Brown \& Levine (2007) employed difference-based kernel estimators. Making use of the local adaptivity of wavelet methods, Cai \& Wang (2008) improved upon previous variance estimation methods using a wavelet thresholding approach on first order differences. Here we present an approach that estimates both the mean and variance accurately by incorporating ASH into a novel wavelet denoising framework. [mention the lack of available software?]

The second task is to denoise a Poisson distributed signal. This often occurs in the experimental sciences such as gamma-ray burst signals in astronomy [references], and more recently in genetics where high throughput sequencing data is of interest. This problem is interesting because there are many distinct ways to perform signal recovery. Variance stabilizing techniques together with normal approximations have been proposed by Donoho (1993) and Fryzlewicz \& Nason (2001), by exploiting the mean-variance relationship of a Poisson distribution. Kolaczyk (1997, 1999a) derived thresholds achieving optimal asymptotic properties in the context of wavelet transformations, similar to the thresholds in the i.i.d. Gaussian case.  However, variance stabilizing methods are computationally inefficient due to the presence of external cycle-spinning, and threshold-based methods may not result in satisfactory finite sample performance. Furthermore, both these methods are quite sensitive to the choice of the primary resolution level as our simulations will show. Multiscale analysis using recursive dyadic partitions within a Bayesian framework was later developed by Kolaczyk (1999b) to make use of a particular form of likelihood factorization, but a relatively inflexible conjugate prior was chosen. We will improve upon the prior (and likelihood) specification by using ASH as the main shrinkage procedure in a Bayesian framework similar to that of Kolaczyk (1999b).

Both the aforementioned problems are harder to deal with than the classical problem with i.i.d. Gaussian errors, and usually requires extension of the ideas from the latter task. Before proceeding to describe the two problems in detail, we will first describe briefly the portion of ASH that performs shrinkage.

As a generic shrinkage method, ASH takes as input a vector of estimates $\hat{\Gb}_i,i=1,...,n$ and their standard errors $\hat{\s}_i\equiv se(\hat{\Gb}_i),i=1,...n$, and outputs shrunken estimates of $\bm{\Gb}$. Specifically, the likelihoods for the true parameters $\Gb_i,i=1,...n$ are assumed to be i.i.d. $N(\Gb_i;\hat{\Gb}_i,\hat{\s}_i^2)$. ASH then assumes exchangeability of the $\Gb_i$'s and sets the following ``shrinkage'' prior on $\Gb_i$ for each $i$:
\begin{eqnarray}
\Gb_i|\bm{\pi}=\sum_{k=1}^m \pi_k N(\Gb_i;0,s_k^2)
\end{eqnarray}
Note that $\bm{\pi}$ and $\bm{s}^2$ are shared across all the $\Gb_i$'s, allowing one to borrow information across all observations. The full likelihood for $\bm{\pi}$ given $\bm{\hat{\Gb}}$ and $\bm{\hat{\s}}$ is then a product of likelihoods for $\bm{\pi}$ given each $\hat{\Gb}_i$ and $\hat{\s}_i$. Since the likelihoods are Gaussian and the priors are mixtures of Gaussians, the posterior distribution of $\Gb_i$ given $\hat{\Gb}_i$ and $\bm{\hat{\pi}}$ is analytically tractable, where the mixing proportions $\bm{\hat{\pi}}$ are maximum likelihood estimates. The posterior mean of $\Gb_i$ is taken to be the point estimate of $\Gb_i$, resulting in shrinkage.

The flexibility and computational efficiency of ASH is immediately obvious, which motivates us to extend this method to the more complex wavelet settings described here. As an Empirical Bayes procedure, ASH borrows information across all available data and is extremely data adaptive. At the same time, the key advantage of ASH over other Empirical Bayes procedures such as EbayesThresh [Johnstone \& Silverman (2005)] and the commonly used ``spike-and-slab'' prior (is there a reference for this?) is that variable variances are allowed as input, a feature which latter methods lack. This is also the primary reason we can use ASH as the only shrinkage procedure in both the problems described above. Furthermore, Gaussian mixtures can effectively approximate any unimodal distribution, allowing for an extremely flexible prior. In tackling these two specific problems, we can speed up the entire signal denoising problem via likelihood approximations (discussed in the next section), allowing our method to be computationally efficient without compromising accuracy. Typical wavelet methods require the specification of a primary resolution level as a ``tuning'' parameter, which could substantially influence the accuracy of the method. However, applying ASH to every resolution in the wavelet transformationwe spares us the need for such a parameter. For an in-depth understanding of the original motivations and applications of ASH, see Stephens (??). With the main shrinkage method accounted for, we will proceed to explore the two aforementioned problems in detail. Note that one only needs to supply a vector of estimates $\hat{\Gb}$ and their standard errors se$(\hat{\Gb})$ to ASH to obtain posterior estimates; hence, we will focus on obtaining these estimates and standard errors when describing our methods.
\section{Method}
\subsection{Gaussian denoising with heteroskedastic errors}
Consider the nonparametric regression model with i.i.d. Gaussian errors:
\begin{eqnarray}\label{eq:1d gaussian model}
Y_i=\mu_i+\Ge_i
\end{eqnarray}
for $i=1,...,n$, where $\bm{Y}$ is the vector of observations, $\bm{\mu}$ is the mean curve, and $\Ge_i$ are independent $N(0,\s^2)$ noise. Assume that $n=2^J$ for some integer $J$, as is standard in the wavelet literature. To motivate our method, we first describe wavelet shrinkage from a Bayesian perspective.

The wavelet coefficients $\bm{d}$ are given by
\begin{eqnarray}
\bm{d}=W\bm{Y}
\end{eqnarray}
where $W$ denotes the orthogonal $n\times n$ matrix corresponding to the orthonormal wavelet basis chosen.
Since $\mathbb{V}(Y)=\s^2I$,
\begin{eqnarray}\label{eq:waveletcoef}
\bm{d}\sim N_n(\bm{\Ga},\s^2I)
\end{eqnarray}
where $\bm{\Ga}=W\mathbb{E}(\bm{Y})$. This implies that the likelihood for $\bm{\Ga}$ factorizes into a product of likelihoods for $\Ga_i$, $i=1,...,n$. As such, a natural and computationally convenient approach is to set independent priors on $\Ga_i$ as a form of ``shrinkage''. Here we apply the priors from ASH to each resolution level separately, resulting in
%Although there are many choices for the prior, we use the Gaussian mixture prior from ASH in our method, given by
\begin{eqnarray}\label{eq:ashprior}
\Ga_{jk}=\sum_{l=1}^m \pi_l^{(j)} N(0,(s_l^{(j)})^2)
\end{eqnarray}
with $\sum_l \pi_l^{(j)}=1$, for $j=1,...,J$. $\pi_i^{(j)}$ and $(s_i^{(j)})^2$ are hyperparameters that are shared between coefficients in the same resolution level, and $m$ is the number of mixture components. In ASH, $m$ is usually chosen to reasonably approximate any unimodal distribution, while maintaining an acceptable computational speed. Note that, for notational convenience, we have switched to a double index following standard wavelet convention for indices: here $j=1,...,J$ is the resolution level, and $k=0,...,2^j-1$ is the location within each resolution level $j$. By applying the inverse wavelet transform to the posterior mean $\hat{\bm{\Ga}}$ that ASH produces, one can obtain the posterior mean of $\bm{\mu}$, which serves as an estimate that minimizes the MSE. One could also construct credible bands using the posterior variances.

While many existing methods perform well in the presence of i.i.d. (Gaussian) errors, heteroskedastic (but still independent) errors present a different challenge. One key obstacle when dealing with heteroskedastic errors is that the likelihood for $\bm{\Ga}$ does not necessarily factorize. Given the true variances and hence the likelihood, one could compute the full posterior distribution using some specified prior. However, this would be computationally cumbersome, and may be infeasible when extended to multiple signals or higher dimensions. In this case, a suitable prior might also be difficult to find. As such, one key aspect of our approach is to treat the wavelet coefficients as if they were independent, so that the true likelihood is approximated by a composite likelihood (see Silverman (1999) for an example where this is done).

To describe our approach in more detail, first assume that the true variance function is known and given by $\mathbb{V}(\Ge_i)=\s_i$, and that the true mean function is given by $\bm{\mu}=(\mu_1,...,\mu_t)$, where $T=2^J$ for some positive integer $J$. Hence
\begin{eqnarray}
\bm{Y}\sim N_n(\bm{\mu},\Sigma)
\end{eqnarray}
where $\Sigma=\textrm{diag}(\s_1,...,\s_T)$ is a diagonal matrix. For the rest of this subsection, we will use the non-decimated wavelet transform (NDWT) instead of the standard DWT to achieve translation invariance. This in turn reduces the presence of artifacts, which often occur near discontinuities in the underlying signal (see eg. Coifman \& Donoho (1995)). Then,
\begin{eqnarray}
\bm{d}\sim N_n(\bm{\Ga},\tilde{\Sigma})
\end{eqnarray}
where $\bm{d}$ is the vector of detail coefficients, $\bm{\Ga}=W\bm{\mu}$, and $\tilde{\Sigma}=W\Sigma W^T$, where $W$ is the matrix associated with the NDWT for a given wavelet basis. By treating the likelihood for $\bm{\Ga}$ as if it were independent, it can be written as follows(using the double index mentioned above):
\begin{eqnarray}\label{eq:likelihood}
L(\bm{\Ga}|\bm{d})=\prod_{j=0}^J\prod_{k=0}^{T-1}P(d_{jk}|\Ga_{jk})
\end{eqnarray}
where $P(d_{jk}|\Ga_{jk})=\phi(d_{jk};\Ga_{jk},\tilde{\Sigma}_{(jk,jk)})$. Note that there are $n$ coefficients at each resolution level instead of $2^j$ for resolution $j$ due to the NDWT. Here $\phi$ denotes the Gaussian density function. Since $\Sigma$ is diagonal, it is easy to see that $\tilde{\Sigma}_{(jk,jk)}=\sum_{i=1}^T \Sigma_{ii}W_{jk,i}^2$. In our method, we use ASH to assign independent priors to $\Ga_{jk}$ as with \eqref{eq:ashprior}. %ASH also allows us to estimate the $\pi_i$'s via a maximum likelihood approach and fix the $\s_i$'s by using a grid of values. This particular choice of prior specification is extremely flexible, being able to approximate any symmetrical zero-centered distribution, as well as allowing for a relatively easy estimation of the hyperparameters through empirical Bayes. See ASH for more details regarding this choice of prior (?)
While ASH produces shrunken estimates of $\Ga_{jk}$ for $j,k>0$, we estimate $\Ga_{00}$ using the corresponding scaling coefficient, which seems intuitively appealing. Finally, we can obtain an estimate of $\mu$ by using the average basis inverse, which is essentially an average of the inverse wavelet transforms for every shift of the data (see Coifman \& Donoho (1995)) since we are using the NDWT. Although the method described here uses a matrix formulation for easier conceptual understanding, the actual NDWT and inverse transform are done through Mallat's pyramid algorithm, taking only $n\log(n)$ time.

Though we have assumed that the true variances were known for mean estimation, the problem of variance estimation itself is a non-trivial one. Here we make the reasonable assumption that the variance function is also spatially structured, in addition to the mean function. The approach we took is similar in spirit to that of Cai \& Wang (2008), making use of wavelet decomposition as well. However, while they use first order differences, we look at the squared residuals. As such, we need a reasonably accurate estimator of the mean function to form sensible residuals. For clarity of presentation, we will first assume that the true mean function is known. Define
\begin{eqnarray}\label{eq:varobs1}
Z_i^2=(Y_i-\mu_i)^2
\end{eqnarray}
to be the ``observations'' for the unknown variance function. Note that $\mathbb{E}(Z_i^2)=\s_i^2$, so that $Z_i^2$ is unbiased for the true variance at point $i$. At this point we can treat this as yet another mean estimation problem, with the added complication that the variance of $Z_i^2$ has a $\chi^2$ distribution squared. To simplify the problem, we approximate this likelihood by a Gaussian likelihood. Of course, better likelihoods and priors could be used that reflect the skewness in the distributions of the ``observations'', but we found that a normal approximation works reasonably well (especially for smoother variance functions), and has the key advantage of being easy and fast to implement. The variance estimation process is then very similar to one for mean estimation as described above, with a few extra details. The full procedure is outlined in Appendix \ref{app:var estimation}.

Now that we can estimate $\bm{\mu}$ given $\bm{\s}^2$ and vice versa, a natural procedure for jointly estimating $\bm{\mu}$ and $\bm{\s}^2$ is an iterative one. That is, we estimate $\bm{\mu}$ given some initial estimate of $\bm{\s}^2$, then estimate $\bm{\s}^2$ given the previous estimate of $\bm{\mu}$, and iterate the process. In our simulations, we have found that two cycles are usually sufficient to produce relatively accurate estimates. To start off the algorithmm, an initial estimate would be the square of first order differences between adjacent points, as discussed in previous variance estimation papers (eg Cai \& Wang (2008)). This estimator has the property that it is approximately unbiased for the variance at the two corresponding points, provided that the mean and the variance function is smooth enough. To be more specific, the initial variance estimates are defined by

@@

Next we turn to another popular nonparametric regression problem and demonstrate how we could make use of ASH. Specifically, we again assume an underlying signal $\Gl_i$, $i=1,...,n$ with $n=2^J$ for some positive integer $J$. Further assume that each data point $Y_i$, $i=1,...,n$ is realized from a Poisson distribution with mean $\Gl_i$ ie. $Y_i=\textrm{Pois}(\Gl_i)$. Our goal is to recover the true intensity $\bm{\Gl}$ as accurately as possible, given $Y_k$. To make our approach easier to understand and relate it to wavelet-based methods, we first summarize the data in a recursive manner (see Kolaczyk (1999) and Nowak (1998):

@@

for resolution $j=0,...,J-1$ and location $k=0,...,2^j-1$. Hence, we are summing more blocks of observations as we move to coarser levels. Another way to define the $Y$'s is

@@

Kolaczyk (1999) set a conjugate shrinkage prior (a mixture of a point mass at 0.5 and a uniform distribution, similar to the ``spike and slab'' prior for the Gaussian case) for the parameters $\frac{\Gl_{j+1,2k}}{\Gl_{jk}}$ and used a Binomial likelihood. However, we work with the parameters $\Ga_{sl}$ as defined in \eqref{eq:poisson wc} and approximate the logit-Binomial (is there such a thing?) likelihoods by Gaussian likelihoods, and replace the two-component prior in Kolaczyk (1999) by the more flexible prior from ASH. Furthermore, these parameters can be more easily connected to the true wavelet coefficients since they consist of (log) differences rather than ratios.

To describe the approximation process in more detail, we drop the subscripts for notational convenience. One direct approximation of the likelihood for $\Ga$ is the Laplace approximation, which results in the Gaussian likelihood given by $\phi(\Ga;\hat{\Ga}_{MLE},\mathbb{V}(\hat{\Ga}_{MLE}))$. Here $\phi$ is the normal density function, and $\hat{\Ga}_{MLE}$ is the maximum likelihood estimator. We would then estimate $\mathbb{V}(\hat{\Ga}_{MLE})$ by the square of the asymptotic standard deviation for $\hat{\Ga}_{MLE}$. Specifically,
% for some appropriate estimator $\hat{\Ga}$. This choice of parameters for the Gaussian approximation is also appropriate in the sense that it minimizes the Kullback-Leibler divergence between a Gaussian distribution and the likelihood for $\Ga$ (also known as moment matching).
